Biodiversity of Japanese Pteridophytes

### 2019-06-30 reports/japan_pteridos_biodiv.Rmd
Use rbind instead of bind_rows for dataframes
bind_rows drops rownames

```{r load-data}
loadd(world_map)
loadd(occ_data_pteridos)
loadd(occ_data_ferns_north)
loadd(occ_data_ferns)
loadd(alpha_div_pteridos)
loadd(alpha_div_ferns_north)
loadd(alpha_div_ferns_north_south)
```

@@

```{r ses-pd-vs-perc-sex-dip-summary}
# Run linear model of SES of PD vs. percentage of sexual diploid ferns
ses_pd_vs_repro_ferns_mod <- lm(ses_pd ~ percent_sex_dip, ses_pd_vs_repro_ferns)

# Extract R2 and pvalue for mentioning in text.
ses_pd_vs_repro_ferns_r2 <- ses_pd_vs_repro_ferns_mod %>% glance %>% pull(r.squared) %>% round(3)

# P is very very small
ses_pd_vs_repro_ferns_mod %>% glance
```

@@

Somewhat surprisingly, SES of PD shows a negative relationship with the percentage of sexual diploid taxa (Fig. \@ref(fig:ses-pd-vs-perc-sex-dip); *R*^2^ = `r ses_pd_vs_repro_ferns_r2`, *P* << 0.05). We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed. It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.

@@

>>>>

```{r load-data}
loadd(world_map)
loadd(occ_data_pteridos)
loadd(occ_data_ferns_north)
loadd(occ_data_ferns)
loadd(alpha_div_pteridos)
loadd(alpha_div_ferns_north)
loadd(alpha_div_ferns_north_south)
loadd(percent_sex_dip_ferns)
loadd(alpha_div_vs_repro_ferns)
```

@@

```{r ses-pd-vs-perc-sex-dip-summary}
# Run linear model of SES of PD vs. percentage of sexual diploid ferns
alpha_div_vs_repro_ferns_mod <- lm(ses_pd ~ percent_sex_dip, alpha_div_vs_repro_ferns)

# Extract R2 and pvalue for mentioning in text.
alpha_div_vs_repro_ferns_r2 <- alpha_div_vs_repro_ferns_mod %>% glance %>% pull(r.squared) %>% round(3)

# P is very very small
alpha_div_vs_repro_ferns_mod %>% glance
```

@@

Somewhat surprisingly, SES of PD shows a negative relationship with the percentage of sexual diploid taxa (Fig. \@ref(fig:ses-pd-vs-perc-sex-dip); *R*^2^ = `r alpha_div_vs_repro_ferns_r2`, *P* << 0.05). We expected that since many apogamous (non-sexual) taxa have identical or very similar DNA sequences to their sexual relatives that communities with more apogamous taxa would be more clustered and communities with more sexual diploids would be overdispersed. It seems that communities with many non-sexual taxa actually tend to include distantly related taxa rather than sister pairs of sexual and non-sexual species.

@@

```{r sex-dip-ferns-map, fig.cap = "Percentage of sexual diploid fern taxa, excluding hybrids."}
make_diversity_map(
  div_data = percent_sex_dip_ferns, 
  world_map = world_map, 
  occ_data = occ_data_ferns, 
  div_metric = "percent_sex_dip",
  metric_title = "% sex. diploid"
) + scale_fill_scico(palette = "bilbao", na.value="white")
```

```{r ses-pd-vs-perc-sex-dip, fig.cap = "Correlation between percentage of sexual diploid fern taxa (excluding hybrids) and standard effect size (SES) of phylogenetic diversity (PD)."}
ggplot(alpha_div_vs_repro_ferns, aes(percent_sex_dip, ses_pd)) +
  geom_point(alpha = 0.4) +
  geom_smooth(
    method = "lm",
    size = 0.5,
    fill = "pink"
  ) +
  annotate("text", 
           x = Inf, y = Inf, 
           label = glue::glue("italic(R) ^ 2 == {alpha_div_vs_repro_ferns_r2}"),
           parse = TRUE,
           hjust = 1.2,
           vjust = 1.2) +
  labs(
    x = "% sexual diploid taxa",
    y = "SES PD"
  ) +
  jntools::standard_theme()
```

```{r mntd-ferns-ns, fig.cap = "Standard effect size (SES) of MNTD of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
make_diversity_map(
    div_data = alpha_div_ferns_north_south, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "mntd.obs.z", 
    metric_title = "MNTD"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns_north_south, mntd.obs.z, "abs", 3), 
      get_limit(alpha_div_ferns_north_south, mntd.obs.z, "abs", 3)
    ))
```

```{r mpd-ferns-ns, fig.cap = "Standard effect size (SES) of MPD of native Japanese fern taxa north of 30.1 degrees latitude, excluding hybrids."}
make_diversity_map(
    div_data = alpha_div_ferns_north_south, 
    world_map = world_map, 
    occ_data = occ_data_ferns, 
    div_metric = "mpd.obs.z", 
    metric_title = "mpd"
  ) + scale_fill_scico(
    palette = "vik", 
    na.value="white",
    limits = c(
      -get_limit(alpha_div_ferns_north_south, mpd.obs.z, "abs", 3), 
      get_limit(alpha_div_ferns_north_south, mpd.obs.z, "abs", 3)
    ))
```

```{r mntd-vs-perc-sex-dip, fig.cap = "Correlation between percentage of sexual diploid fern taxa (excluding hybrids) and MNTD."}

# Run linear model of SES of PD vs. percentage of sexual diploid ferns
mtnd_vs_percent_sex_mod <- lm(mntd.obs.z ~ percent_sex_dip, alpha_div_vs_repro_ferns)

# Extract R2 and pvalue for mentioning in text.
mtnd_vs_percent_sex_mod_r2 <- mtnd_vs_percent_sex_mod %>% glance %>% pull(r.squared) %>% round(3)

ggplot(alpha_div_vs_repro_ferns, aes(percent_sex_dip, mntd.obs.z)) +
  geom_point(alpha = 0.4) +
  geom_smooth(
    method = "lm",
    size = 0.5,
    fill = "pink"
  ) +
  annotate("text",
           x = Inf, y = Inf,
           label = glue::glue("italic(R) ^ 2 == {mtnd_vs_percent_sex_mod_r2}"),
           parse = TRUE,
           hjust = 1.2,
           vjust = 1.2) +
  labs(
    x = "% sexual diploid taxa",
    y = "mntd.obs.z"
  ) +
  jntools::standard_theme()
```

```{r mntd-vs-el, fig.cap = "Correlation between elevation and MNTD."}

# Run linear model of SES of PD vs. percentage of sexual diploid ferns
mtnd_vs_el_mod <- lm(mntd.obs.z ~ elevation, alpha_div_vs_repro_ferns)

# Extract R2 and pvalue for mentioning in text.
mtnd_vs_el_r2 <- mtnd_vs_el_mod %>% glance %>% pull(r.squared) %>% round(3)

ggplot(alpha_div_vs_repro_ferns, aes(elevation, mntd.obs.z)) +
  geom_point(alpha = 0.4) +
  geom_smooth(
    method = "lm",
    size = 0.5,
    fill = "pink"
  ) +
  annotate("text",
           x = Inf, y = Inf,
           label = glue::glue("italic(R) ^ 2 == {mtnd_vs_el_r2}"),
           parse = TRUE,
           hjust = 1.2,
           vjust = 1.2) +
  labs(
    x = "Elevation (m)",
    y = "mntd.obs.z"
  ) +
  jntools::standard_theme()
```

```{r mpd-vs-el, fig.cap = "Correlation between elevation and mpd."}

# Run linear model of SES of PD vs. percentage of sexual diploid ferns
mtnd_vs_el_mod <- lm(mpd.obs.z ~ elevation, alpha_div_vs_repro_ferns)

# Extract R2 and pvalue for mentioning in text.
mtnd_vs_el_r2 <- mtnd_vs_el_mod %>% glance %>% pull(r.squared) %>% round(3)

ggplot(alpha_div_vs_repro_ferns, aes(elevation, mpd.obs.z)) +
  geom_point(alpha = 0.4) +
  geom_smooth(
    method = "lm",
    size = 0.5,
    fill = "pink"
  ) +
  annotate("text",
           x = Inf, y = Inf,
           label = glue::glue("italic(R) ^ 2 == {mtnd_vs_el_r2}"),
           parse = TRUE,
           hjust = 1.2,
           vjust = 1.2) +
  labs(
    x = "Elevation (m)",
    y = "mpd.obs.z"
  ) +
  jntools::standard_theme()
```
